---
title: "Example Births R Markdown File"
author: "Nick Brazeau"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    highlight: tango
    theme: lumen
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = T, warning = F, message = F, error = F, eval = T)

# These are my global options, I can set these to change the default behavior of my R code chunks
# Note the include = FALSE means it won't be included in my report but will be evaluated 

```

```{r libraries}
# data wrangling
library(tidyverse)

# table outputs
library(knitr) # and so much more
library(kableExtra)

# fancy outputs
library(DT)
library(stargazer)
library(plotly)


```

# Reproducibiltiy Example
```{r}
# source("~/Desktop/ex.R")
# df %>% 
#   dplyr::group_by(favoriteprogram) %>% 
#   dplyr::summarise(n=n()) %>% 
#   ggplot(data=.) + 
#   geom_bar(aes(x=favoriteprogram, y=n), stat="identity")

```

# Births Dataset
Previously, the R Instructors made me do a bunch of work cleaning and recoding this database. I tidied it and then saved it as an RDS object using the script `01_tidy-data.R` in the `analysis` directory of my **Births R Project**. I am now going to read in my tidy dataset and investigate if prenatal care reduces the risk of preterm birth. 
```{r, echo=F}
# this will not show my code
births <- readRDS("../data/births_afterhw2.RDS")
```
```{r, echo=T}
# note the difference here, this WILL show my code in the report
births <- readRDS("../data/births_afterhw2.RDS")
```
## Simply for Speed (and check out eval)
> This is a great time to point out that I subsetted the births dataset to 500 observations for speed. I did this with the `set.seed()` function. My code looked like this:     

```{r, echo=T, eval=F}
set.seed(43)
births <- births[sample(1:nrow(births), size=500, replace = F), ]
write_rds(x=births, path="data/births_afterhw2.RDS")
```

# Code Chunks
## Toggling Results
#### Hide
```{r, eval = T, results='hide'}

summary(births$mrace)

```
#### ASIS 
```{r, eval=T, results='asis'}

# some of this was set in my global options but as a reminder

summary(births$mrace)

```

^ Some code will not output in "pretty" formats. 

##### Prettier Table Outpet
Ways around this but data.frames or tibbles to knitr ["kables"](https://rdrr.io/cran/knitr/man/kable.html) privdes a nice option. [Kableextra](https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html) has additional features. 

#### Hide
```{r, echo=T, eval=F}

births %>% 
  dplyr::select(mage, raceeth_f) %>% 
  dplyr::group_by(raceeth_f) %>% 
  dplyr::summarise(meanage=mean(mage)) %>% 
  knitr::kable(.)

```
#### ASIS 
```{r, echo=F, eval=T, results='asis'}

births %>% 
  dplyr::select(mage, raceeth_f) %>% 
  dplyr::group_by(raceeth_f) %>% 
  dplyr::summarise(meanage=mean(mage))%>% 
  knitr::kable(.)

```

#### Kableextra asis  
```{r, echo=F, eval=T, results='asis'}


births %>% 
  dplyr::select(mage, raceeth_f) %>% 
  dplyr::group_by(raceeth_f) %>% 
  dplyr::summarise(meanage=mean(mage)) %>% 
  mutate_if(is.numeric, round, 2) %>% # remember mutate if?
  knitr::kable(., "html") %>%  # note must tell kableExtra html or latex
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F, position = "center")

```



## Controlling Figure Size 
### Reasonable
```{r, results='asis', fig.align='center', fig.height=6, fig.width=6}

births %>%
  dplyr::group_by(mage) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(data=.) + geom_bar(aes(x=mage, y=n), stat="identity")
```

### Ridiculous
```{r, results='asis', fig.align='right', fig.height=1, fig.width=6}

births %>%
  dplyr::group_by(mage) %>% 
  dplyr::summarise(n=n()) %>% 
  ggplot(data=.) + geom_bar(aes(x=mage, y=n), stat="identity")
```

# Outputs
## Data

### Raw
```{r}
DT::datatable(births, extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 5,
              dom = 'Bfrtip', 
              buttons = c('csv')))
```
### Summarized 
```{r}
births %>% 
  dplyr::group_by(mage) %>% 
  dplyr::summarise(n=n()) %>% 
  DT::datatable(., extensions='Buttons',
               options = list(
              searching = T,
              pageLength = 5,
              dom = 'Bfrtip', 
              buttons = c('csv')))
```

## Bivariate Analyses 
### TableOne
```{r}
devtools::install_github("kaz-yos/tableone") # used in HW2, need for github for kableone functipn from PR #32 (https://github.com/kaz-yos/tableone/pull/32)
library(tableone)

vars_of_interest = c("pnc5_f", "preterm_f", "smoker_f", "wksgest", "sex", "mage", "meduc", "raceeth_f", "weeknum", "include_allpass")
tableone::kableone(tableone::CreateTableOne(data=births[, vars_of_interest]))

```

## Regression Outputs
```{r}
births <- births %>% 
  mutate(pnc5_f = factor(pnc5_f, levels = c("No Early PNC", "Early PNC"))) %>% 
  mutate(preterm_f = factor(preterm_f, levels=c("term", "preterm")))

crude <- glm(preterm_f ~ pnc5_f, 
            data = births,
            family = binomial(link = "logit"))

mage <- glm(preterm_f ~ pnc5_f + mage + poly(mage, 2, raw = T),
            data = births,
            family = binomial(link = "logit"))

magesm <- glm(preterm_f ~ pnc5_f + mage + poly(mage, 2, raw = T) + smoker_f, 
            data = births,
            family = binomial(link = "logit"))
```

### Broom
```{r}
broom::tidy(mage, exponentiate=TRUE, conf.int=TRUE)
broom::tidy(mage)
broom::confint_tidy(mage)
exp(broom::confint_tidy(mage))
exp(confint(mage))

```

### Stargazer
```{r, results='asis'}
stargazer::stargazer(crude, mage, magesm, title = "Results of Some Births Regression",
          dep.var.labels=c("Preterm"),
          add.lines=list(c("AIC", round(AIC(crude),1), round(AIC(mage),1), round(AIC(magesm),1))),
          no.space=F,
          ci=TRUE, ci.level=0.95,
          single.row=F,
          align=T,
          type = "html")


# stargazer::stargazer(crude, mage, magesm, title = "Results of Some Births Regression", 
#           dep.var.labels=c("Preterm"),
#           add.lines=list(c("AIC", round(AIC(crude),1), round(AIC(mage),1), round(AIC(magesm),1))),
#           no.space=F,
#           ci=TRUE, ci.level=0.95, 
#           single.row=F,
#           align=T, type = "latex")

```

# Graphics
## plotly 
```{r}
library(plotly)
plotObj <- ggplot(data=births, aes(x=mage, y=wksgest, color=preterm_f)) +
  geom_point()

plotly::ggplotly(plotObj)

```


# Other Engines
### Using Bash
```{bash}
echo "Hi Nick" | sed 's/Nick/Sara/'
```

### Using Python
```{python}
def printsnake( str ):
   print "SSSSSS" + str
   return;
   
printsnake("hello")
   
```


